---
title: 'Final Assignment: Predicting Director Compensation'
author: "Jasper Ginn (s6100848)"
date: "April 8, 2019"
output:
  pdf_document: 
    number_sections: false
bibliography: bib.bib
geometry: margin=1.7cm
header-includes:
 \usepackage{float}
 \restylefloat{table}
 \usepackage{setspace}
 \usepackage{xcolor}
 \usepackage[font={small}]{caption}
 \singlespacing
 \floatplacement{figure}{H}
---
<!-- Set font size -->
\fontsize{11}{22}

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.pos="H", fig.width = 7, fig.height = 4, echo = FALSE, 
                      fig.align="center", warning = FALSE, message = FALSE, eval=FALSE)

rm(list=ls())

# Load blm package
library(blm)

# Load ggplot
library(ggplot2)
library(gridExtra)
library(dplyr)

# Get the directors data
data("directors")

# remove where outcome / covariate age is NA
# (blm cannot handle missing values)
directors_with_na <- directors
directors <- directors %>%
  filter(!is.na(Age),
         !is.na(Compensation))

# Keep for later
dirs <- directors

# Preprocess directors data
# This is a subset of the original (proprietary) dataset
# 5 sectors have not been included due to bimodality in the data. 
# See the folder 'data-raw' for the script used to create the directors data.
library(dplyr)
directors <- directors %>%
  # Log compensation
  mutate(Compensation = log(Compensation)) %>%
  # Subtract mean from Age variable
  mutate(Age = Age - mean(Age),
         Male = as.numeric(Male) - mean(as.numeric(Male)))

# Load models 
# i.e. prevent R from re-running the models the entire time when knitting the doc (this takes a long time).
# Instead, run code line-by-line if you need to.
dirmod2 <- readRDS("models/model2.rds")
resm2 <- readRDS("models/model3.rds")
```

```{r julia setup}
# Set up blm julia
blm_setup() 
```

# Introduction

The gender gap in compensation remains a pervasive problem despite efforts to achieve income parity between men and women in the private and public sector. In 2019, the difference in compensation between men and women stood at an estimated median 17\% in the general workforce [CITE], and it appears that the glass ceiling may stretch all the way to the top. In 2016, a report found that the average growth in compensation for male execute board members of S&P500 companies^[The S\&P500 is a stock market index that tracks the 500 most profitable companies listed on the US stock exchange.] was twice that of female executive board members [CITE]. 

The boards of American companies consists of *executive* and *independent* directors. Execute directors are employed by the firm and usually have high positions within the company (for example the Chief Executive Officer), while independent directors are concerned primarily with oversight [CITE]. Most of the literature on director compensation concerns executive directors rather than independent directors [CITE]. In this paper, I examine the gender gap in compensation among $336$ independent directors in three sectors and $52$ companies using Bayesian Linear Regression.

The results indicate that we cannot conclude definitively that male independent directors receive a higher remuneration than female independent directors after controlling for age and sector. Model fit statistics and evaluation suggest that linear regression is not an appropriate method to analyse these data because director compensation is highly correlated within companies.  

The paper is structured as follows. Section 2 elaborates on the differences between Frequentist and Bayesian statistical inference. In turn, section 1 outlines the hypotheses, model specification and reports the results of the analyses. Section three concludes.

# Differences in Bayesian and Frequentist Inference

For a large part, differences between Frequentist and Bayesian inference stems from their respective view of uncertainty, how this is captured by probabilities and, in particular, what *causes* this uncertainty. In the Frequentist framework, a probability is defined as the number of times an event will occur in the long run (@miller2014john, p.21). In other words, it is the limiting value of $m$ successes in a sequence of $n$ trials for a particular event, or:

$$
\tag{1}
\begin{aligned}
p = \lim_{n \to \infty} \frac{m}{n}
\end{aligned}
$$

This definition has two important implications. The first is that talking about probability makes sense only in the context of infinite trials. The second implication is that probability converges to some *fixed* quantity as the number of trials go infinity, and our uncertainty about the true value of this quantity reduces as we repeatedly take more (or larger) samples. Hence, this definition implies that the only source of randomness by which our estimate $\hat{p}$ differs from the true value $p$ comes from the data [CITE], which may differ from sample to sample due to, for example, sampling error.

The Bayesian framework considers probabilities as a means of quantifying uncertainty about knowledge (@gelman2013bayesian, pp. 11-13). Even though the 'true' parameter value may be fixed, we are limited by our knowledge of this value. Hence, the uncertainty by which we make statements about the world changes as we collect more information which is represented by the posterior distribution, which may be looked upon very loosely as the collection 'true' values conditional on how certain we are about the veracity of this knowledge.

## Objective and Subjective Knowledge

By the very definition of probability in the Frequentist framework, evidence can originate only from the data. This is not the case in Bayesian statistical inference, where inferences are based on a mix of domain expertise (*prior* or *belief*) and evidence from the data. This makes perfect sense in the Bayesian framework; if we are certain about our knowledge then we can constrain the parameter space by injecting what we know *a priori*. Hence, a Bayesian looks upon prior beliefs as just another source of knowledge that has been translated into a probability density. 

```{r calculate prior variance effect}
# Posterior shrinkage
dirmod3 <- blm("Compensation ~ Sector + Age + Male",
                data=directors) %>%
  # Need to delete posterior when we change settings
  # (only has an effect if posterior already present)
  delete_posterior() %>%
  # Set an informative prior on the coefficients
  # Male (b4) is speculated to have 17% pay increase over women
  set_prior("b4", mu = .17, sd=0.001) %>%
  # Update sampling settings
  set_sampling_options(chains = 2, iterations = 20000,
                       burn = 4000, thinning=3) %>%
  # Set initial values
  # If we draw from the priors the starting values will be too large
  # This is an issue for MH because it takes many more iterations to converge
  set_initial_values(chain_1 = list("b" = c(4, 2, 3, 8, 10), "sigma"= 1),
                     chain_2 = list("b" = c(2, 4, 7, 2, 3), "sigma"=2)) %>%
  # Sample the posterior distribution
  sample_posterior()

# Change prior and run for another 600000 observations
dirmod3 <- dirmod3 %>%
  set_prior("b4", mu = .17, sd=0.01) %>%
  update_posterior(iterations=20000)
# Same
dirmod3 <- dirmod3 %>%
  set_prior("b4", mu = .17, sd=0.1) %>%
  update_posterior(iterations=20000)

# Save
saveRDS(dirmod3, "data/model_effects_prior.rds")
```

```{r effect of prior variance, fig.height=2, fig.width=7, fig.cap="Effect of adjusting the prior variance of the coefficient for variable Male (a) on the posterior distribution (b). The mean is set at M=.17 which represents a 17 percent increase in compensation for males versus females. Reducing the prior variance represents increased certainty about the estimate of our domain knowledge and weights the information from the data less strongly compared to domain knowledge. The dashed lines indicate the value of the posterior means.", eval=FALSE}

dirmod3 <- readRDS("data/model_effects_prior.rds")
# Retrieve posterior samples for male variable, burn and label by prior variance 
library(purrr)
pd <- get_value(dirmod3, "posterior") %>%
  get_value("samples") %>%
  map(function(x) {
    r1 <- x[4001:20000,"Male"]
    r2 <- x[24001:40000, "Male"]
    r3 <- x[44001:60000, "Male"]
    return(
      data.frame(
        "Male" = c(r1, r2, r3),
        "Prior" = factor(c(rep(1, length(r1)),rep(2, length(r2)), rep(3, length(r3))),
                         labels = c("0.001", "0.01", "0.1"))
      )
    )
  }) %>%
  bind_rows()

# Calculate SD by prior
pd_sd <- pd %>%
  group_by(Prior) %>%
  summarize(psd = sd(Male),
            pmn = mean(Male))

# Posterior z-value
zscore <- function(mean, sd, theta) (mean - theta) / sd
z1 <- zscore(pd_sd$pmn[1], pd_sd$psd[1], .17)
z2 <- zscore(pd_sd$pmn[2], pd_sd$psd[2], .17)
z3 <- zscore(pd_sd$pmn[3], pd_sd$psd[3], .17)

# Posterior scaling of sd
ps1 <- pd_sd$psd[1]^2 / as.numeric(levels(pd_sd$Prior)[1])^2 # 841
ps2 <- pd_sd$psd[2]^2 / as.numeric(levels(pd_sd$Prior)[2])^2 # 35
ps3 <- pd_sd$psd[3]^2 / as.numeric(levels(pd_sd$Prior)[3])^2 # 0.53

# Posterior means
pms <- round(pd_sd$pmn, digits=3)

library(ggplot2)
library(ggpubr)
priorvar <- ggplot(data.frame(
  "Male" = c(rnorm(10000, mean=.17, sd=0.001), 
             rnorm(10000, mean=.17, sd=.01), 
             rnorm(10000, mean=.17, sd=.1)),
  "Prior" = factor(c(rep(1, 10000), rep(2, 10000), rep(3, 10000)), labels = c("0.001", "0.01", "0.1"))
  ),
  aes(x=Male, fill=Prior)) +
  geom_histogram(alpha=0.4, position="identity", color="black") +
  theme_blm(text_size = 9) +
  theme(axis.title = element_blank(),
        axis.text.y=element_blank()) +
  scale_fill_brewer(palette = "Set1",
                    name="Prior variance") +
  labs(title = "(a) Prior distribution")

# Posterior distribution
postvar <- ggplot(pd, aes(x=Male, fill=Prior)) +
  geom_histogram(alpha=0.4, position="identity", color="black") +
  theme_blm(text_size = 9) +
  theme(axis.title = element_blank(),
        axis.text.y=element_blank()) +
  geom_vline(xintercept = pms[1], color=RColorBrewer::brewer.pal(3, "Set1")[1], 
             linetype="dashed", size=1.2) +
  geom_vline(xintercept = pms[2], color=RColorBrewer::brewer.pal(3, "Set1")[2], 
             linetype="dashed", size=1.2) +
  geom_vline(xintercept = pms[3], color=RColorBrewer::brewer.pal(3, "Set1")[3], 
             linetype="dashed", size=1.2) +
  scale_fill_brewer(palette = "Set1") +
  labs(title= "(b) Posterior distribution")

# Arrange plots with common legend
ggarrange(priorvar, postvar, ncol=2, common.legend=TRUE, legend="top")
```

The claim that the Frequentist approach is more objective would hold only if the data collection process, (pre-)processing steps and analysis are guaranteed to be objective. This is a tenuous assumption at best in the social sciences; statistics is fraught with subjective decisions during data collection, manipulation and analysis @berger1988statistical. Hence, the critique leveled at Bayesians boils down to the practice of incorporating domain knowledge *explicitly* through the use of a prior. It should be noted that, if an analyst chooses to use priors such that they represent a complete lack of knowledge (for example, a uniform distribution), results obtained using Frequentist and Bayesian methods often yield comparable results.

## Methods of Estimation and Hypothesis Testing

Whether a statistician regards the data or the parameters as a random variable determines their choice of estimation method. A Frequentist, believing that variation originates from data, will want to optimize the 'likelihood' of the data conditional on the parameters. That is, we find the *most likely combination* of parameters $\hat{\theta}$ that explain the data and that provide consistent and asymptotically unbiased estimators, or:

$$
\tag{1}
\hat{\theta} = \arg \max_{\theta \in \Theta} \text{Log Likelihood}(\theta |\text{data}) 
$$

Conversely, given that a Bayesian thinks of the data as fixed and the parameters as random variables, they are interested in finding the distribution of the parameters and hence the source of uncertainty in our beliefs after seeing the evidence:

$$
\tag{2}
p(\theta | \text{data}) = \frac{p(\text{data}|\theta) p(\theta)}{p(\text{data})}
$$

The question of how the definition of probability and the methods of estimation impact inference under these frameworks is illustrated by the difference in interpretation of the confidence interval and the credible interval. When we calculate the confidence interval, the boundaries of the confidence intervals are interpreted as random variables due to sampling error since they estimate the frequency of sampling means. Therefore, we should interpret it as ... With a credible interval, we do not have this source of variability, which gives rise to the definition that the parameter is contained in the credible interval with some probability because the parameter space is assumed to be known under the assumptions by which we arrived at the posterior distribution.

Further implications are to be found in the way we test hypotheses in these frameworks. In the Frequentist framework, we usually partition parameter space into an acceptance and a rejection region based on some null and alternative hypothesis. On the basis of a test statistic, computed from the data, we then decide whether or not the result we observe is likely to occur due to chance. We use either confidence intervals or $p$-values to express support for the alternative hypothesis, but, importantly, we can never quantify support in favor of the null hypothesis. This is not so in the Bayesian Framework. Given that the data can be viewed as a means to update a prior belief, hypothesis testing is a relative statement about the degree to which the evidence found in the data supports this initial belief or contradicts it. 

<!--
TYPE I & II errors, advantages of bayes factor, not effect sizes
-->

## Predictions under the model

Bayesian models yield the posterior probability distribution of the model parameters. This allows us to produce a posterior predictive distribution for each subject in the data set. Unlike Frequentist model results, such distributions can be transformed almost endlessly, and it provides us with a useful way to understand whether the model adheres to its core assumptions by means of *posterior predictive checks* (PPC). PPC require us to simulate data under the model and to compare the simulated outcome variable to the observed outcome variable, which results in a proportion sometimes called the Bayesian p-value. If we desire to interpret this p-value like we do the traditional, Fisherian p-value, then we must assume that, under the assumption that a particular test is not violated, the Bayesian p-value is uniformly distributed. Otherwise, we cannot interpret it as a proportion. We often find that this assumption does not hold true for posterior predictive checks. To this end, we can simulate data under various conditions to observe how the statistic behaves. This is illustrated in figure XX for two posterior predictive checks included in the R library blm.

```{r ppc simulations, fig.cap="Distributions of posterior predictive p-values for 1.000 simulated data sets. In plots (a) and (b), the simulated data are drawn from a normal without any violations of the linear regression assumptions. In plot (c), the assumption of homoskedasticity is violated in each of the simulations. In plot (d), the assumption of independence of errors is violated in each of the simulations. The color indicates the severity of the violation; the green bars indicate mild violation, blue indicates medium violation and red indicates severe violation. The script used to generate the data and run the simulations can be found on GitHub.", fig.width=7, fig.height=3, eval=TRUE}

# Convenience function to flatten array
flatten_array <- function(x) {
  
  # To list and return
  lapply(1:dim(x)[3], function(y) {
    ss <- x[,,y]
    colnames(ss) <- c("normality", "heteroskedasticity", "independence")
    # To df
    ss <- as.data.frame(ss)
    # Return
    return(ss)
  })
  
}

# Read data
normal <- readRDS("data/simulations/ppc_normal_data.rds")
finsim <- readRDS("data/simulations/final.rds")

normal <- finsim$noviolation$results %>%
  flatten_array() %>%
  bind_rows()
heterosked <- finsim$heterosked$results %>% 
  flatten_array() %>%
  bind_rows(.id="severity") %>%
  mutate(severity = factor(severity, labels = finsim$heterosked$degrees %>% names()))
indep <- finsim$indep$results %>% 
  flatten_array() %>%
  bind_rows(.id="severity") %>%
  mutate(severity = factor(severity, labels = finsim$heterosked$degrees %>% names()))

# PPC heteroskedasticity, no violation
p1 <- ggplot(data.frame(x=normal[,2]), aes(x=x)) +
  geom_histogram(color="black", fill="lightgrey") +
  scale_x_continuous(limits=c(0,1)) + 
  theme_blm(text_size = 10) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank()) +
  labs(title = "(a) PPC heteroskedasticity", subtitle="      no violations")
# PPC independence, no violation
p2 <- ggplot(data.frame(x=normal[,3]), aes(x=x)) +
  geom_histogram(color="black", fill="lightgrey") +
  scale_x_continuous(limits=c(0,1)) + 
  theme_blm(text_size = 10) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank()) +
  labs(title = "(b) PPC independence", subtitle="      no violations")
# PPC heteroskedasticity, assumption violated
p3 <- ggplot(heterosked, aes(x=heteroskedasticity, fill=severity)) +
  geom_histogram(color="black", alpha=0.4, position = "identity")  + 
  theme_blm(text_size = 10) + 
  scale_fill_brewer(palette = "Set1", direction=-1) +
  scale_x_continuous(limits=c(-0.02,1)) +
  theme(axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      legend.position = "none") +
  labs(title = "(c) PPC heteroskedasticity", subtitle="      assumption violated")
# PPC independence, violated
p4 <- ggplot(indep, aes(x=independence, fill=severity)) +
  geom_histogram(color="black", alpha=0.4, position = "identity")  + 
  scale_fill_brewer(palette = "Set1", direction=-1) +
  scale_x_continuous(limits=c(-0.02,1)) +
  theme_blm(text_size = 10) +
    theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none") +
  labs(title = "(d) PPC independence", subtitle="      assumption violated")

# Merge plots
grid.arrange(p1, p2, p3, p4)
```

# Methods & Results

Table XX below shows the descriptive statistics for each variable. The outcome variable **compensation** is given in thousands of Great British pounds and has been log-transformed. The variable **Age** has been grand-mean centered to facilitate interpretation and estimation.

```{r descriptive statistics, eval=FALSE, results='asis'}
out <- capture.output(stargazer::stargazer(directors_with_na, summary=TRUE, median=TRUE))
cat(paste0(out[-1:-3], "\n"))
```

The model is given by the following equation:

$$
\tag{1}
\begin{aligned}
\hat{\log(\text{compensation}_i)} = 
&\beta_0 + \beta_{1} [\text{age}_i - \bar{\text{age}}] + \beta_2\text{male}_i + \beta_3\text{SectorServices}_i + \beta_4\text{SectorBasicMaterials}_i + \epsilon_i
\end{aligned}
$$

Given that we measure the compensation on a log scale, the coefficients we derive from the model must be interpreted as change in percentages. This forces us to rethink the priors we set on the model. Additionally, even if we know very little of the effect of age on compensation (or gender for that matter), we can set some reasonable assumptions in terms of upper and lower bounds. For age, we now assume that, as age increases, compensation increases as well. But, given that we are not certain about the extent to which it increases, we set the standard deviation of this estimate to $.2$, representing a spread of approximately $20\%$^[For values close to $0$, $exp(x) \approx 1 + x$]. For gender, we know from earlier studies that the gender gap is $17\%$. However, theory suggests that the gender gap is lower or non-existent at top-tier firms. Hence, we set this prior a mean of $.1$ with a spread of $.1$, which reflects our uncertainty of the estimate. 

In accordance with the theory, we hypothesize that there is at most a small difference in compensation between male and female directors compared to the rest of the work force, which we will set at $5\%$.

$$
\text{H}_1\text{: } \beta_{\text{Male}} \leq .05
$$

Furthermore, we hypothesize compensation only increases with age. That is:

$$
\text{H}_2\text{: } \beta_{\text{Age}} > 0 
$$

Finally, we specify the unconstrained hypothesis as:

$$
\text{H}_u\text{: } \beta_\text{Male}, \beta_{\text{Age}}
$$

<!--
These hypotheses will be included in the models we construct. (Hoitjtink p.24).

PRA is the weight of a particular Bayes factor compared to the other hypotheses. That is, it is a proportion of the total evidence.

PRB is the probability of the marginal likelihood of each hypothesis, or $P(H_i | model)$. In essence, this functions as an indication of error probabilities; if we 'cut' the parameter space into many pieces, so does our confidence in our hypothesis go down.

The quantities PRB are the Bayesian error probabilities 

BF sensitive to outliers (Hoijtink pp.31-32) ==> check for outliers before BF. Also violation of model assumptions using ppc. We should use a robust BF but we don't have that. Anyway, interpret the BF with a grain of salt.

Bayes' Factor is sensitive to the choice of prior ==> something about the size of complexity. (Hoijtink pp.29-30)
-->

```{r model1, include = FALSE}
# Model 2: Age, gender and sector -----

# We have three sectors: financial (base case), services and Basic materials 

# Create the model
dirmod2 <- blm("Compensation ~ Sector + Age + Male",
                data=directors) 

# Retrieve parameter names
dirmod2 %>% get_parameter_names()

# Update the model specification
dirmod2 <- dirmod2 %>%
  # Need to delete posterior when we change settings
  # (only has an effect if posterior already present)
  delete_posterior() %>%
  # Change the uninformative priors for the intercept and the sectors
  # We are reasoning in percentage change now. So uninformative priors should reflect this
  # on the same scale. (except the intercept)
  set_prior("b0", mu = 0, sd = 100) %>%
  set_prior("b1", mu = 0, sd = log(1000)) %>%
  set_prior("b2", mu = 0, sd = log(1000)) %>%
  # Set an informative prior on the coefficients
  # Age centered at 0 with .2 variance
  set_prior("b3", mu = 0, sd = .2) %>%
  # Males (b4) are speculated to have 17% pay increase over women
  set_prior("b4", mu = .1, sd=.1) %>%
  # Change the sampler of Age to metropolis hastings (random walk)
  # Lambda parameter controls the variance of the (normal) proposal distribution
  set_sampler("b3", type="MH", lambda=0.01) %>%
  # Update sampling settings
  set_sampling_options(chains = 2, iterations = 80000,
                       burn = 15000, thinning=5) %>%
  # Set initial values
  # If we draw from the priors the starting values will be too large
  # This is an issue for MH because it takes many more iterations to converge
  set_initial_values(chain_1 = list("b" = runif(5, 0, 10), "sigma"=1),
                     chain_2 = list("b" = runif(5, 0, 10), "sigma"=2)) %>%
  # Compute null model (the intercept-only model)
  compute_null_model() %>%
  # Set hypotheses
  set_hypothesis("H1", "b4 < .05") %>% # Gender: less than 5 percent difference
  set_hypothesis("H2", "b3 > 0") %>%   # Age: larger than 0
  # Sample the posterior distribution
  sample_posterior()

# Print model specification
print(dirmod2)

# Evaluate the accepted draws
dirmod2 %>% 
  evaluate_accepted_draws() # ~ 40-45%

# Look at burn-in diagnostics
dirmod2 %>% 
  evaluate_convergence_diagnostics() # GR: chains have converged

# Check history
plot(dirmod2, "history") # This seems OK
# Check autocorrelation
plot(dirmod2, "autocorrelation") # OK

# Evaluate the effect of MH on the effective sample size
dirmod2 %>% 
  evaluate_effective_sample_size() # OK

# View densities
plot(dirmod2, "density") # OK

# Did the null model converge?
plot(dirmod2, "nullmodel") # OK

# Summary
summary(dirmod2)
# ==> Male not a predictor
# ==> age a predictor
# ==> Both sectors increase compensation for directors

# R-squared value
# Compute a z-score difference from the posterior to the prior
# Idea: how much does the posterior shift when confronted with new data?
#        calculate on the scale of the posterior
thprior <- rnorm(1, mean=dirmod2$priors$b4$mu, sd=dirmod2$priors$b4$sd)
ps <- get_posterior_samples(dirmod2) 
pmu <- apply(ps, 2, mean)
psd <- apply(ps, 2, sd)
z <- abs((pmu["Male"] - thprior) / psd["Male"])
shrink <- psd[5] / .01 # Factor of 6 ==> this means that the posterior variance is 6 times larger than the prior variance.

# Plot
plot(dirmod2$rsq)
median(dirmod2$rsq$rsquared) # R-squared ~ .089
quantile(dirmod2$rsq$rsquared, c(0.025, 0.25, 0.5, 0.75, 0.975)) # 95% CCI [.045, .15]

# Posterior predictive checks
dirmod2 <- dirmod2 %>%
  evaluate_ppc(p=1)

# Residual plot
errs <- data.frame(
  "residuals_model" = resid(dirmod2),
  "predicted_model" = predict(dirmod2),
  "company" = as.character(directors$Company[!is.na(directors$Compensation)])
)

library(ggplot2)
# Color residuals by company
p1 <- ggplot(errs, aes(x=predicted_model, y=residuals_model, color=as.factor(company))) +
  geom_point() +
  theme_blm() +
  theme(legend.position="none")
library(purrr)
# Correlation within each board
rc <- errs %>%
  split(.$company) %>%
  map(function(x) {
    x$residuals_lagged <- c(NA, x$residuals_model[1:length(x$residuals_model)-1])
    cor(x$residuals_model, x$residuals_lagged, use="complete.obs")
  }) %>%
  unlist()

rc <- data.frame(
  "cor" = unname(rc),
  "comp" = names(rc)
)
# Plot correlation within each board
p2 <- ggplot(rc, aes(x=reorder(comp, -abs(cor)), y=cor)) +
  geom_bar(stat="identity") +
  theme_blm() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) 
# ==> correlation by boards, should do a multilevel model here to accomodate hierarchical structure of the data.

library(gridExtra)
grid.arrange(p1, p2, ncol=2)
# Summary
summary(dirmod2$ppc) # Violating independence assumption
p <- plot(dirmod2$ppc, "normality") ; plot(p) # Two-sided test ==> violation ==> less skewed compared to model 1
p <- plot(dirmod2$ppc, "heteroskedasticity") ; plot(p) # One-sided test ==> violation (data are more heteroskedastic) ==> more extreme than model 1
p <- plot(dirmod2$ppc, "independence") ; plot(p) # Two-sided test ==> violation (data are correlated) ==> less extreme than model 1 so we're better taking into account group effects

## Autocorrelation ==> not strange. Board members of the same COMPANY are more alike than board members of other company. Hence the correlation in the errors.

# Model fit
dirmod2 %>%
  get_value("DIC") %>%
  summary()# ==> DIC is 487.895 with 6 parameters 
dirmod2 %>%
  get_value("null_model") %>%
  get_value("DIC") %>%
  summary() # ==> 513 so the model above fits significantly better

# Save this model
saveRDS(dirmod2, "models/model2.rds")
```

```{r PPC results, eval=FALSE, fig.width=7, fig.height=2, fig.cap="A random subset of observed and simulated correlation coefficients for model 1 (left) and model 2 (right). The figure shows that the observed residuals (in red) are much more correlated than is reasonable under the model (represented by the blue, simulated values)."}
# Collect ppc results
ppc_results <- dirmod2$ppc$data$independence
# Mean correlation over samples
mcor <- mean(ppc_results[,2])

# Get values for independence ppc

# Model 1
indeppc2 <- dirmod2 %>% get_value("ppc") %>%
  get_value("data") %>% .$independence %>%
  as.data.frame() %>%
  tidyr::gather(type, value) %>%
  mutate(type = ifelse(type == "V1", "Simulated", "Observed"),
         samp = runif(1:n())) %>%
  filter(samp <= .15) %>%
  group_by(type) %>%
  mutate(index = 1:n()) %>%
  ggplot(aes(x=index, y=value, color=type)) +
    geom_point() +
    theme_blm() +
    scale_color_brewer(palette = "Set1") +
    theme(axis.text.x = element_blank(),
          axis.title.y = element_blank(),
          legend.title = element_text(rel(1.2)),
          legend.text = element_text(rel(1.2)),
          axis.title.x = element_blank(),
          legend.position = "none") +
  labs(title="(b) Model 2", subtitle = "PPC independence of errors")
# Plot
indeppc2
```

The model results are presented in table XX. The convergence diagnostics show that the posterior distributions have forgotten their initial states and have converged to their stable distributions; there is little to no trace of autocorrelation and the Gelman-Rubin statistic is close to $1$, which indicates that the chains have converged to the same stationary distribution. The DIC for the model of interest (model 2) is $488$ compared to $512$ for the intercept-only model, which indicates it yields a better fit. The MC error, which should be less than $5\%$ of the standard deviation, is negligible.

From the posterior predictive checks (table XY), we observe that the data are not independent $(p_{\text{independence}} = 0$. Directors who are in the same board tend to are more similar to each other than directors from other boards. This is not unexpected: the data are hierarchical in nature such that individuals are nested in boards. Indeed, the results of the random effects model shows that the intra-class correlation coefficient $\rho$ equals $57\%$, meaning that the expected correlation of two randomly picked directors from the same company is $\hat{r}=.573$.^[The results of running all different stages of a multilevel model are presented in another document.] 

The model results indicate that gender is not a predictor of director compensation $(\beta_{\text{male}} = .062; 95\% \ \ \text{CCI} = [-.08, .205])$, and credible interval for this coefficient tells us that there is a lot of uncertainty in this estimate. Age has a small, positive effect on director compensation $(\beta_{\text{age}} = .01; 95\% \ \ \text{CCI} = [.002, .018])$. When holding the other variables constant, every one-year increase above the mean age leads to a $1\%$ increase in compensation. Finally, it appears that directors in the sectors 'basic materials' and 'services' earn some $25\%$ and $34\%$ more than their counterparts in the financial sector. The mean $R^2$ value for this model is $R^2 = .09; 95\% \ \ \text{CCI} = [.041, .15])$ [CITE GELMAN], indicating that we are explaining between $4\%$ and $15\%$ of the variance in the data.

```{r MLM intercept-only model}
library(rjags)

# Reload directors
data("directors")
# Remove missing values
directors <- directors %>%
  filter(!is.na(Age),
         !is.na(Compensation)) %>%
  # Refactor
  mutate(Company = factor(as.character(Company)))

# Preprocess directors data
library(dplyr)
directors <- directors %>%
  # Log compensation
  mutate(Compensation = log(Compensation),
         Male = as.numeric(Male) - 1) %>%
  # Create level 2 variables
  group_by(Company) %>%
  mutate(avgAge = mean(Age, na.rm=TRUE),
         propMale = sum(Male) / n(),
         # Group-mean center individual data
         Age = Age - mean(Age),
         Male = Male - mean(Male)) %>%
  # Ungroup
  ungroup() %>%
  # Grand-mean center group variables
  mutate(avgAge = avgAge - mean(avgAge),
         propMale = propMale - mean(propMale)) %>%
  # Sort data (makes it easier later)
  arrange(Sector)
  
# Jags data
dir_jags <- with(directors, list(## Outcome for individuals (level 1)
  
                                 compensation = Compensation,
                                 # Age of individuals
                                 age=Age,
                                 # Gender of individuals
                                 gender=Male,
                                 
                                 ## Company-level variables (level 2)
                                 
                                 # Company indicator
                                 company=as.numeric(as.factor(Company)),
                                 # Average age of directors for each company
                                 avgAge=avgAge,
                                 # Proportion of males for each company
                                 avgMale=propMale,
                                 
                                 ## Group totals
                                 
                                 # Number of individuals
                                 n=nrow(directors),  
                                 # Number of companies
                                 k=length(unique(Company)),
                                 
                                 ## Number of predictors
                                 
                                 # Number of individual-level predictors
                                 p1=2,
                                 # Number of company-level predictors
                                 p2=2))

# Intercept-only model
io_mod <- "model {

  ### Level 2 ==> companies

	# Priors (companies)
	tau_u0 ~ dgamma(.01, .01)
	# Hyperprior for mean
  gamma_00 ~ dnorm(0, 1.0E-4)
  # For each company
	for (j in 1:k) {
	  # For each company, draw from normal
    b0[j] ~ dnorm(gamma_00, tau_u0)
	}

	### Level 1 ==> individuals

  # Priors
  # Individual precision
	tau ~ dgamma(.01, .01)  # standard deviation of fixed effect (variance within companies)
  # For each individual
	for (i in 1:n) {
		compensation[i] ~ dnorm(mu[i], tau) # Combine likelihood and priors
		mu[i] <- b0[company[i]] # Linear combination for each person
	}

  # Invert gamma
  sigma_e <- 1 / sqrt(tau)
  sigma_u0 <- 1 / sqrt(tau_u0)

}"

# Run the model in JAGS
# Initial values
dir_inits <- list(
  init1 <- list(tau=runif(1), tau_u0=runif(1)),
  init2 <- list(tau=runif(1), tau_u0=runif(1))
)

# Specify model in JAGS
mod_io <- jags.model(textConnection(io_mod), 
                     data = dir_jags,
                     inits = dir_inits,
                     n.chains=2)

# Burn
update(mod_io, n.iter=60000)

# Draw samples
params <- c("sigma_e", "sigma_u0", "gamma_00")
# Run the chain
resm1 <- coda.samples(mod_io, variable.names = params, n.iter=500000, thin = 10)

# Convergence?
plot(resm1) # Ok
gelman.plot(resm1) # Ok

# DIC
DICm1 <- dic.samples(mod_io, thin=5, n.iter=20000, type="pD") # 288.6

# MAP values
MAPm1 <- apply(do.call(rbind.data.frame, resm1), 2, mean)

# Coerce to matrix
resmat <- do.call(rbind, resm1)

# New matrix with same dims
ve <- matrix(0, ncol=1, nrow=nrow(resmat))

# FOr each row
# Posterior distribution of ICC
for(i in seq_along(1:nrow(resmat))) {
  ve[i,] <- c(resmat[i,3]^2 / sum(resmat[i,-1]^2))
}

quantile(ve, c(0.025, .25, .5, .75, .975))
# VE between .45 and .69

# Save the model and DIC
saveRDS(resm1, "models/ML_interceptonly.rds")
saveRDS(DICm1, "models/ML_io_DIC.rds")
```

Given the hierarchical nature of the data, we next run a linear mixed effects model. This model corresponds to the following equation.

$$
\tag{3}
\begin{aligned}
\text{compensation}_{ij} =
&\gamma_{00} + u_{0j} + \gamma_{10}\cdot \text{Age}_{ij} + e_{ij}
\end{aligned}
$$

Where $\gamma_{00}$ is the overall intercept and $u_{0j}$ is a company-specific error term. Notice that the fit of this model is much better than that of the previous models ($DIC=280$), indicating that the multilevel approach seems appropriate. The marginal and conditional R-squared values (@nakagawa2013general). [The marginal R-squared takes into account only the fixed part of the variance, while the conditional R-squared takes into account the fixed and random parts] are $R^2_{\text{M}} = .0056; \ 95\% \ \text{CCI}=[.0008, .015]$ and $R^2_{\text{C}} = .5398; \ 95\% \ \text{CCI}=[.4807, .6]$ respectively. This indicates that the fixed part of the model (age) explains almost no variation in the data, but the fixed and random parts together explain some $54\%$ of the total variation. Hence, we can be confident that director compensation is best modelled using company-specific intercepts, but that we do not have the right variables at either level 1 or 2 that would help us explain the heterogeneity in compensation among directors within boards and between boards.

```{r MLM level-1 predictors}
# Initial values
dir_inits <- list(
  init1 <- list(tau=runif(1), tau_u0=runif(1)),
  init2 <- list(tau=runif(1), tau_u0=runif(1))
)

# Model
mlm_2 <- "model {

  ### Level 2 ==> companies

	# Priors (companies)
	tau_u0 ~ dgamma(.01, .01)
	# Hyperprior for mean
  gamma_00 ~ dnorm(0, 1.0E-4)
  gamma_10 ~ dnorm(0, 1.0E-4)
  # For each company
	for (j in 1:k) {
	  # For each company, draw from normal
    b0[j] ~ dnorm(gamma_00, tau_u0)
	}

	### Level 1 ==> individuals

  # Priors
  # Individual precision
	tau ~ dgamma(.01, .01)  # standard deviation of fixed effect
  # For each individual
	for (i in 1:n) {
		compensation[i] ~ dnorm(mu[i], tau) # Combine likelihood and priors
		mu[i] <- b0[company[i]] + # Intercept
		         gamma_10*age[i] #+ gamma_20*gender[i] # Level 1 variables
	}

  # Invert gamma
  sigma_e <- 1 / sqrt(tau)
  sigma_u0 <- 1 / sqrt(tau_u0)

}"

# Specify model in JAGS
mod_io <- jags.model(textConnection(mlm_2), 
                     data = dir_jags,
                     inits = dir_inits,
                     n.chains=2)

# Burn
update(mod_io, n.iter=60000)

# Draw samples
params <- c("sigma_e", "sigma_u0", "gamma_00", "gamma_10")
# Run the chain
resm2 <- coda.samples(mod_io, variable.names = params, n.iter=500000, thin = 10)

# DIC
DICm2 <- dic.samples(mod_io, thin=5, n.iter=20000, type="pD") # 280.9 => +- 9 difference

# Calculate variance explained
# MAP values for model 1
MAPm1 <- apply(do.call(rbind.data.frame, resm1), 2, mean)
MAPm2 <- apply(do.call(rbind.data.frame, resm2), 2, mean)

# Map values
#MAPve <- apply(ve_2, 2, median)

# Save model
saveRDS(DICm2, "models/ML_final_DIC.rds")
saveRDS(resm2, "models/model3.rds")
```

```{r calculate variance explained for mixed model, eval=TRUE}
# Calculate R^2. Marginal (only explained by predictors) & Conditional (predicted by variance component)
# See: https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.2041-210x.2012.00261.x

# Function to calculate marginal & conditional R^2
# @param X design matrix containing fixed part of the model
# @param w coefficient matrix
# @e_1 random part of level 1
# @e_2 random part of level 2
MLM_R2 <- function(X, w, e_1, e_2) {
  
  # Component 1: fixed effect
  e_f <- var(X %*% w)

  # Compute marginal & conditional R^2
  mar <- (e_f) / (e_f + e_1 + e_2) # Variance explained by fixed factors (marginal R^2)
  con <- (e_f + e_2) / (e_f + e_1 + e_2) # Variance explained by fixed and random factors (conditional R^2)
  
  # Return
  return(
    c(mar, con)
  )
  
}

# Coerce to matrix
resmat <- do.call(rbind, resm2)

# New matrix with same dims
ve_2 <- matrix(0, ncol=2, nrow=nrow(resmat))

# X for fixed effect
X <- matrix(directors$Age, ncol=1)

# FOr each row
# Posterior distribution of variance explained (pseudo R-squared)
for(i in seq_along(1:nrow(resmat))) {
  ve_2[i,] <- MLM_R2(X, 
                 resmat[i,2],
                 resmat[i,3],
                 resmat[i,4])
}

# Median var explained
mve <- apply(ve_2, 2, median)
```

```{r r-squared plot, fig.cap="Bayesian R-squared value for model 1 (a) and model 2 (b). The proportion of cases in which the R-squared value of model 2 exceeds that of model 1 is .99. Figure (c) shows the marginal (blue) and conditional (red) R-squared values for the posterior distribution. The marginal R-squared indicates the amount of variance explained by the fixed part of the model; the conditional R-squared indicates the amount of variance explained by the fixed and random part of the model.", fig.height=2, fig.width=7, eval=TRUE}
library(latex2exp)
library(ggplot2)
# Plot r-squared
# Model 1
p1 <- dirmod2 %>% get_value("rsq") %>% plot() + theme_blm(text_size = 8) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        plot.subtitle = element_blank()) +
  labs(title = latex2exp::TeX("\\textbf{(b) Bayesian $R^2$ posterior (model 2)}"))
# Model 2 (multilevel)
veg <- ve_2 %>%
  as.data.frame() %>%
  rename("MR2" = V1, "CR2" = V2) %>%
  #select(CR2) %>%
  tidyr::gather(type, rsq)
# Plot
p2 <-  ggplot(veg, aes(x=rsq, fill=type)) +
    geom_histogram(alpha=.4, color="black") +
    theme_blm(text_size=8) +
    scale_fill_brewer(palette="Set1") +
    geom_vline(xintercept=veg %>% filter(type=="MR2") %>% 
                 select(rsq) %>% pull() %>% 
                 median(), linetype="dashed") +
    geom_vline(xintercept=veg %>% filter(type=="CR2") %>% 
                 select(rsq) %>% pull() %>% 
                 median(), linetype="dashed") +
    theme(axis.title = element_blank(),
          axis.text.y = element_blank(),
          legend.position = "none") +
    labs(title = latex2exp::TeX("\\textbf{(c) Marginal and Conditional $R^2$ posterior}"))

# Grid
grid.arrange(p1, p2, ncol=2)
```

\begin{table}[H] \centering 
\begin{small}
  \caption{Model results } 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & \multicolumn{3}{c}{Compensation (GBR '000, logged)} \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\\[-1.8ex] & Intercept-only & Full model & Linear mixed effects \\
\\[-1.8ex] & (blm) & (blm) & (JAGS) \\
\hline \\[-1.8ex] 
\textit{(a) Fixed} & & & \\
Constant & 4.991 (4.935, 5.048) & 4.820 (4.731, 4.901) & 4.993 (4.875, 5.112) \\ 
SectorBasic Materials &  & .225 (.082, .37) &  \\ 
SectorServices &  & .294 (.172, .415) &  \\ 
Male & & .062 ($-$.08, .205) & \\ 
Age & & .010 (.002, .018) & .009 (.003, .015) \\ 
\hline \\[-1.8ex] 
\textit{(b) Random} & & & \\
$\sigma^2_e$ & .52 & .5 & .342 (.315, .372) \\
$\sigma^2_{u0}$ & & & .399 (.322, .504) \\
\hline \\[-1.8ex] 
\textit{(c) Model Fit} \\
Observations & 336 & 336 & 336 \\
Companies & & & 52 \\
DIC & 513 & 488 & 281 \\ 
Penalty. & 2 & 6 & 49 \\ 
R$^{2}$ & 0 & .09 & .006 (M), .536 (C) \\ 
\hline \\[-1.8ex] 
\textit{(d) Post. Pred. Checks} & & & \\
Normality &  & .357 &  \\
Homoskedasticity &  & .612 & \\
Independence &  & 0 & \\
\hline \\[-1.8ex] 
\textit{(e) Bayes' Factors (model 2 only)} & & & \\
Hypothesis & BF (complexity, fit) & PMPa & PMPb \\
\hline \\[-1.8ex] 
$\text{H}_1: \beta_\text{Male} \leq .05$ & 4294.5 (.812, 1) & .382 & .291  \\
$\text{H}_2: \beta_{\text{Age}} > 0$ & 129.8 (.498, .992) & .618 & .472 \\
$\text{H}_u: \beta_{\text{Male}}, \beta_{\text{Age}}$ &  &  & .237 \\
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{Baseline is sector 'Financials' for models (1) and (2).} \\
\end{tabular} 
\end{small}
\end{table}

```{r classical residuals, eval=FALSE}
plot(predict(dirmod2), resid(dirmod2))
```



# References
